Spin canonical divisors of log stacky curves

### 2015-06-30 stacky_write_up/spin_cring.tex
edits finished for monster theorem

\begin{align*}
	\initial_\prec(I) &= \initial_\prec(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] 
										 + \langle U_i: 1 \le i \le n-1 \rangle
										 + \langle V \rangle,\\
\end{align*}
where 
$V=\{x_i y_j: 1\le i\le m, 1\le j\le n\}$
 and $U_i$ is the set of monomials of the form $\prod_{j=1}^{i} y_j^{a_j}$ with $a_j\in \mathbb{Z}_{\ge 0}$ such that
\begin{enumerate}
	\item[(1)] $\sum_{j=1}^i a_j c_j \ge c_{i+1},$\\
	\item[(2)] there does not exist $(b_1, \ldots b_i)\ne (a_1, \ldots a_i)$ with all $b_j\le a_j$ and $\sum_{j=1}^i b_jc_j\ge c_{i+1},$ \\
	\item[(3)] there does not exist $h<i$ such that $\sum_{j=1}^h y_j^{a_j}\ge c_{h+1}$
\end{enumerate}

@@

{\bf Part (b):}
Define $U_i$ and $V$ as in the lemma's statement and define
\[
	T = (\bigcup_{i=1}^{n-1} U_i) \cup V.
\]

We check that $S$, $T$, and $\prec$ meet the hypothesis of Lemma \ref{lem:relations_from_generators_induction}.  We have already shown $S$ forms a $k$-basis for $R_\halfcan$ over $R_{\halfcan'}$ giving us condition \ref{lem:relations_from_generators_induction}(1).  Our choice of order for monomials in $k[y_1, \ldots, y_n]$ and block order for those in $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ implies that $T\succ S\succ k[u, x_1, \ldots, x_m]$ giving us condition \ref{lem:relations_from_generators_induction}(2).

Now, suppose $f\in k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ is a monomial not contained in $k[u, x_1, \ldots, x_m]$, meaning there is some $j$ such that $y_j|f$.  Further suppose $f\not\in \langle V \rangle$, meaning that for each $i\in \{1, \ldots m\}$, $x_iy_j\not |f$ so $x_i\not|f$.  Therefore, $f\in k[u, y_1, \ldots, y_n]$.  We note that $S$ generates $y_j (k[u,y_1, \ldots, y_n])$ as a $k$-algebra.  That is, all monomials of $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ are contained in
\[
	k[u, x_1, \ldots, x_m] \cup V \cup \{y_1, \ldots, y_n\} k[u, y_1, \ldots, y_n]. 
\]
\todo{is $\{y_1, \ldots, y_n\} k[u, y_1, \ldots, y_n]$ appropriate notation to mean all possible products between some $y_j\in \{y_1, \ldots, y_n\}$ and some element in $k[u, y_1, \ldots, y_n]$?}
Thus, we see that if $f\in k[u, y_1, \ldots, y_n]$, then $f=u^b \prod_{j=1}^l {y_j}^{a_j}\in S$ for some $1\le l\le n$, $b\in \mathbb{Z}_{\ge 0}$, and $a_j\in \mathbb{Z}_{\ge 0}$ for all $j$ with $a_l> 0$.  For $j\in \mathbb{N}$ with $l < j\le n$, assign $a_j=0$.  If $i\in \{1, \ldots, n\}$, then $y_i f$ is either in $S$ or there is some $h\in \mathbb{N}$ such that $i\le h\le \max(i,l)$ satisfying $\prod_{j=1}^h y_j^{a_j}\not\in S_{\sum_{j=1}^h c_j a_j}$ \todo{Peter: terrible subscripts...}, and for all $r<h$, we have $\prod_{j=1}^r {y_j}^{a_j}\in S_{\sum_{j=1}^r c_ja_j}$.  We see that this latter case when $y_if\not\in S$ is equivalent to $\prod_{j=1}^h {y_j}^{a_j}\in U_h$.  Thus, since $S$ generates $\{y_1, \ldots, y_n\} k[y_1, \ldots, y_n]$ and $\bigcup_{i=1}^{n-1} U_i$ generates every monomial in $\{y_1, \ldots, y_n\} k[y_1, \ldots, y_n]-S$, we have shown that all monomials of $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ are contained in
\begin{align*}
				& k[u, x_1, \ldots, x_m] \cup \langle V \rangle \cup S \cup \langle \bigcup_{i=1}^{n-1} U_i \rangle \\
	\subseteq \; 	& S \cup \langle T\rangle \cup \initial_\prec(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \; \cup \; k[u, x_1, \ldots, x_m].
\end{align*}
This gives us condition \ref{lem:relations_from_generators_induction}(3).  Thus, the conditions of Lemma \ref{lem:relations_from_generators_induction} are met, and by applying its conclusion we prove (b) of this lemma.  \todo{Peter: Checking that all the conditions of Lemma \ref{lem:relations_from_generators_induction}(2) are met takes a lot of space and clutter.  On the other hand, it is somewhat important.  Should it be included?}

>>>>

\begin{align*}
	\initial_\prec(I) &= \initial_\prec(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] 
										 + \langle U_i: 1 \le i \le n-1 \rangle
										 + \langle V \rangle,\\
\end{align*}
where 
$V=\{x_i y_j: 1\le i\le m, 1\le j\le n\}$
 and $U_i$ is the set of monomials of the form $\prod_{j=1}^{i} y_j^{a_j}$ with $a_j\in \mathbb{Z}_{\ge 0}$ such that
\begin{enumerate}
	\item[(1)] $\sum_{j=1}^i a_j c_j \ge c_{i+1},$\\
	\item[(2)] there does not exist $(b_1, \ldots b_i)\ne (a_1, \ldots a_i)$ with all $b_j\le a_j$ and $\sum_{j=1}^i b_jc_j\ge c_{i+1},$ \\
	\item[(3)] there does not exist $h<i$ such that $\sum_{j=1}^h y_j^{a_j}\ge c_{h+1}$.
\end{enumerate}

@@

{\bf Part (b):}
Define $U_i$ and $V$ as in the lemma's statement and define
\[
	T = \left(\bigcup_{i=1}^{n-1} U_i \right) \cup V.
\]

We check that $S$, $T$, and $\prec$ meet the hypothesis of Lemma \ref{lem:relations_from_generators_induction}.  In part (a), we have shown $S$ \todo{What is $S$. I presume it is the complement, but can you spell it out?} forms a $k$-basis for $R_\halfcan$ over $R_{\halfcan'}$ implying condition (1) of Lemma ~\ref{lem:relations_from_generators_induction} holds.  Our choice of order for monomials in $k[y_1, \ldots, y_n]$ and block order for those in $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ implies that $T\succ S\succ k[u, x_1, \ldots, x_m]$, implying that condition (2) of Lemma \ref{lem:relations_from_generators_induction} holds.

It only remains to check condition (3) of Lemma ~\ref{lem:relations_from_generators_induction}. To do this, suppose $f\in k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ is a monomial not contained in $k[u, x_1, \ldots, x_m]$, meaning there is some $j$ such that $y_j|f$.  Further suppose $f\not\in \langle V \rangle$, meaning that for each $i\in \{1, \ldots m\}$, $x_iy_j \nmid f$. Since $y_j \mid f$ but $x_iy_j \nmid f$, we obtain $x_i\nmid f$.  Therefore, $f\in k[u, y_1, \ldots, y_n]$.  We note that $S$ generates $y_j \cdot (k[u,y_1, \ldots, y_n])$ as a $k$-algebra.  That is, all monomials of $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ are contained in
\[
	k[u, x_1, \ldots, x_m] \cup V \cup \left(\cup_{i=1}^n y_i \cdot k[u, y_1, \ldots, y_n]\right). 
\]
Thus, we see that if $f\in k[u, y_1, \ldots, y_n]$, then $f=u^b \prod_{j=1}^l {y_j}^{a_j}\in S$ for some $1\le l\le n$, $b\in \mathbb{Z}_{\ge 0}$, so that $a_l > 0$ and $a_j\in \mathbb{Z}_{\ge 0}$ for $1 \leq j \leq l$. For $j\in \mathbb{N}$ with $l < j\le n$, assign $a_j=0$. \todo{Is there a reason why you don't just take the product up to $n$ and let $l$ be the highest number so that $a_l \neq 0$? This seems simpler to me}
If $i\in \{1, \ldots, n\}$, then $y_i f$ is either in $S$ or there is some $h\in \mathbb{N}$ such that $i\le h\le \max(i,l)$ satisfying $\prod_{j=1}^h y_j^{a_j}\not\in S_{\sum_{j=1}^h c_j a_j}$ \todo{Peter: terrible subscripts... Aaron: Indeed, why not just omit the subscript?}, 
and for all $r<h$, we have $\prod_{j=1}^r {y_j}^{a_j}\in S_{\sum_{j=1}^r c_ja_j}$.  We see that this latter case when $y_i f \notin S$ is equivalent to $\prod_{j=1}^h {y_j}^{a_j}\in U_h$. \todo{Aaron: Why to we see this? Presumably this is using conditions 1,2,3 but it would be helpful for you to spell out which. Also, what are the latter and former cases? I don't see the casework anywhere}
Thus, since $S$ generates $\cup_{i=1}^n y_i \cdot k[u, y_1, \ldots, y_n]$ and $\bigcup_{i=1}^{n-1} U_i$ generates every monomial in $\{y_1, \ldots, y_n\} k[y_1, \ldots, y_n]-S$, we have shown that all monomials of $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ are contained in
\begin{align*}
				& k[u, x_1, \ldots, x_m] \cup \langle V \rangle \cup S \cup \left\langle \bigcup_{i=1}^{n-1} U_i \right\rangle \\
	\subseteq \; 	& S \cup \langle T\rangle \cup \initial_\prec(I') k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \; \cup \; k[u, x_1, \ldots, x_m].
\end{align*}
This shows condition (3) of Lemma ~\ref{lem:relations_from_generators_induction} holds.  Thus, the conditions of 
Lemma ~\ref{lem:relations_from_generators_induction} are met. Finally, Lemma ~\ref{lem:relations_from_generators_induction} implies part (b).  \todo{Peter: Checking that all the conditions of Lemma \ref{lem:relations_from_generators_induction}(2) are met takes a lot of space and clutter.  On the other hand, it is somewhat important.  Should it be included?}
